\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-19 paper.qmd
Added draft texts for tuning

@@

>>>>

After establishing the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-rmlowdenshex (a)). To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points. This involves calculating the number of points within each hexagon. Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid (see @eq-equationp2). Next, bins with a standard count less than a certain benchmark value are removed (see @fig-rmlowdenshex (b)). The following steps will help in determining a suitable value for removing low-density hexagons:

1. Plot the distribution of the standardized counts (see @fig-stdctsScurve).
2. Examine the distribution of counts.
3. Select the first quantile value if the distribution is skewed.

Furthermore, selecting the benchmark value for removing low-density hexagons is important. Removing unnecessary bins may lead to the formation of long edges and an uneven 2D model. Hence, rather than solely relying on the benchmark value to identify hexagons for removal, it's essential to consider the standard number of points in the neighboring hexagons of the identified low-density ones (see @fig-rmlowdenshex (c)). If neighboring bins also show low counts, only those bins will be removed. The remaining bins are used to construct the 2D model.

The benchmark value for removing low-density hexagons ranges between $0$ and $1$. When analyzing how these benchmark values influence model performance, it's essential to observe the change in MSE as the benchmark value increases (see @fig-diagnosticpltScurvelwd). The MSE shows a gradual increase as the benchmark value progresses from $0$ to $1$. Evaluating this rate of increase is important. If the increment is not considerable, the decision might lean towards retaining low-density hexagons.

@@

Creating a smooth 2D representation (see @fig-modelScurve (c)) requires removing edges that connect distant bin centroids within the triangular mesh (see @fig-modelScurve (b)). These edges only exist in the 2D model and do not extend into higher dimensions, ensuring that their removal does not impact the model in higher dimensions. While specific criteria for determining the benchmark value to remove long edges do not exist, the following steps provide an approach to identifying a suitable threshold:

1. Plot the distribution of the 2D Euclidean distances (see @fig-distScurve).
2. Identify the largest difference between consecutive distance values.
3. Take the largest distance value corresponding to this difference as the benchmark value.

### 2024-05-19 paper.tex
Added draft texts for tuning

@@

>>>>

After establishing the hexagonal grid with an appropriate number of
bins, some hexagonal bins may have few or no data points within them
(see \textbf{?@fig-rmlowdenshex} (a)). To ensure comprehensive coverage
of the NLDR data, it is necessary to select hexagonal bins with a
considerable number of data points. This involves calculating the number
of points within each hexagon. Then, the standard count is computed by
dividing the number of points within each hexagon by the maximum number
of points in the grid (see \textbf{?@eq-equationp2}). Next, bins with a
standard count less than a certain benchmark value are removed (see
\textbf{?@fig-rmlowdenshex} (b)). The following steps will help in
determining a suitable value for removing low-density hexagons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the standardized counts (see
  \textbf{?@fig-stdctsScurve}).
\item
  Examine the distribution of counts.
\item
  Select the first quantile value if the distribution is skewed.
\end{enumerate}

Furthermore, selecting the benchmark value for removing low-density
hexagons is important. Removing unnecessary bins may lead to the
formation of long edges and an uneven 2D model. Hence, rather than
solely relying on the benchmark value to identify hexagons for removal,
it's essential to consider the standard number of points in the
neighboring hexagons of the identified low-density ones (see
\textbf{?@fig-rmlowdenshex} (c)). If neighboring bins also show low
counts, only those bins will be removed. The remaining bins are used to
construct the 2D model.

The benchmark value for removing low-density hexagons ranges between
\(0\) and \(1\). When analyzing how these benchmark values influence
model performance, it's essential to observe the change in MSE as the
benchmark value increases (see \textbf{?@fig-diagnosticpltScurvelwd}).
The MSE shows a gradual increase as the benchmark value progresses from
\(0\) to \(1\). Evaluating this rate of increase is important. If the
increment is not considerable, the decision might lean towards retaining
low-density hexagons.

@@

Creating a smooth 2D representation (see \textbf{?@fig-modelScurve} (c))
requires removing edges that connect distant bin centroids within the
triangular mesh (see \textbf{?@fig-modelScurve} (b)). These edges only
exist in the 2D model and do not extend into higher dimensions, ensuring
that their removal does not impact the model in higher dimensions. While
specific criteria for determining the benchmark value to remove long
edges do not exist, the following steps provide an approach to
identifying a suitable threshold:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the 2D Euclidean distances (see
  \textbf{?@fig-distScurve}).
\item
  Identify the largest difference between consecutive distance values.
\item
  Take the largest distance value corresponding to this difference as
  the benchmark value.
\end{enumerate}
